---
title: "Lab 10: Filtering Data"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
knitr::opts_chunk$set(warning = FALSE)
knitr::opts_chunk$set(message = FALSE)
knitr::opts_chunk$set(fig.height = 5)
knitr::opts_chunk$set(fig.width = 8.5)
knitr::opts_chunk$set(out.width = "100%")
knitr::opts_chunk$set(dpi = 300)

library(readr)
library(ggplot2)
library(dplyr)
library(viridis)
theme_set(theme_minimal())
```

## Pick a dataset

Today you will again have a choice of three datasets to work with. You can
choose either to investigate the daily team ELO ratings in the NBA:

```{r}
nba <- read_csv("https://statsmaths.github.io/stat_data/nba_elo_daily.csv")
```

Metadata from US senator's tweets:

```{r}
tweets <- read_csv("https://statsmaths.github.io/stat_data/senator_tweets_meta.csv")
```

Or swear words given in the movies of Quentin Tarantino:

```{r}
tarantino <- read_csv("https://statsmaths.github.io/stat_data/tarantino.csv")
```

Pick a dataset (feel free to delete the others from the Rmd file if you would
like), and work on the instructions in the next section.

## Instructions

This lab is similar to the last. You'll pick a dataset and then generate an
analysis of that data using data visualizations. The difference here is that
instead of a single plot, I would like you to construct 3-4 plots that
together tell a linear story. Each plot should be separated by a sentence or
two describing what the viewer should take away from the plot. Try to keep the
plots similar in some way; perhaps each simply highlights a different subset
of the data but has the same underlying layers.

Notice that each of the datasets for today are larger than you will probably
be able to use. You will need to filter the data to a particular team, year,
subset of curse words. You may further filter the data to highlight an even
smaller subset of the data within each plot.

## Analysis

Tarantino is known for his dramaticized action scenes and also his rather colorful language. In this project, we will be analyzing the number of curse words that Tarantino uses in a variety of his movies. First, lets just look at the amount of curse words in his movies. 

```{r}
ggplot(tarantino, aes(movie, word)) +
  geom_point() +
  geom_point(aes(color=word), size=2)
```

This isn't the plots for the project but I was curious. 

Now, lets see in what movies he says "fuck" or a variation of "fuck" in his movies!!

```{r}
fucks <- filter(tarantino, word %in% c("fuck", "fucked", "fucker", "fuckers", "fuckface", "fuckhead", "fucking", "fucks", "fuckup"))
```

```{r}
 ggplot(fucks, aes(movie, word)) +
  geom_point() +
  geom_point(aes(color=word), size=2)
```

As you can clearly see, the word "fuck" appeared in all of Tarantino's listed movies. Lets also see how often and when these cuss words appeared. 

```{r}
ggplot(fucks, aes(movie, minutes_in)) +
  geom_point() +
  geom_point(aes(color=word), size=2) +
  xlab("Movie") +
  ylab("Minutes into the Movie") 
```

More specifically, lets see when the word "fuck" is used. 

```{r}
just_fuck <- filter(fucks, word=="fuck")
```

```{r}
ggplot(just_fuck, aes(movie, minutes_in)) +
  geom_point(aes(color=movie), size=2) +
  xlab("Movie") +
  ylab("Minutes into Movie")
```

That is a lot of "fucks." Lets zoom in closer, to see how many times he says it in the first 10 minutes of his movies. 

```{r}
ten_mins <- filter(just_fuck, minutes_in<=10.00)
```

```{r}
ggplot(ten_mins, aes(movie, minutes_in)) +
  geom_point(aes(color=movie), size=2) +
  xlab("Movie") +
  ylab("Minutes into the Movie")
```

Honestly, this is a lot less than what I had expected to see. There were only two movies in which he cursed in the first ten minutes. 